---
title: "Reproducible Research: Peer Assessment 1"
author: "Tijesunimi Adebiyi"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries
```{r echo=FALSE}
library(dplyr)
library(imputeTS)
library(kableExtra)
library(ggplot2)
```


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")

#I want to convert the minute into text and as HHMM
intervalCh <- as.character(activity$interval)

activity$time <-  sapply(intervalCh, function(x){
    strlength <- nchar(x)
    return(paste(paste(rep("0", 4-strlength),collapse =  ""),x, 
                     sep = ""))
    }, USE.NAMES = FALSE)
activity$dateTime <- with(activity,
                          strptime(paste(date,time),
                                   format = "%F %H%M"))
activity$date <- with(activity,
                          strptime(date,
                                   format = "%F"))


```



## What is mean total number of steps taken per day?

```{r}
meanPerDay = activity %>% group_by(date) %>% 
    summarise(mean.steps = mean(steps, na.rm = T),
              median.steps = median(steps, na.rm = T))
hist(meanPerDay$mean.steps,
     main = "Histogram of Mean Steps per day",
     xlab = "Mean Steps per day")
```
This is a table of median and mean per day

```{r}
kbl(meanPerDay) %>% 
    kable_styling()
```

## What is the average daily activity pattern?
```{r}
plot(meanPerDay$date, meanPerDay$mean.steps,  type = "l")
```



## Imputing missing values
```{r}
dfInputMissing <- activity

#Fill missing values in mean with moving averages
inputtedMean <- na_ma(meanPerDay)

#Fill missing values in original dataframe
#with mean of steps in the day

missingRows <- activity[is.na(activity$steps),]

missingInputed <- missingRows
missingInputed$steps <- inputtedMean$mean.steps[
    match(missingInputed$date,meanPerDay$date)
]
dfInputMissing[is.na(dfInputMissing$steps),] <- missingInputed

replacedMeanPerDay <-  dfInputMissing %>% group_by(date) %>% 
    summarise(mean.steps = mean(steps, na.rm = T),
              median.steps = median(steps, na.rm = T))

```

This is the Histogram of the data with missing values replaced
```{r}

hist(replacedMeanPerDay$mean.steps,
     main = "Histogram of Mean Steps per day",
     xlab = "Mean Steps per day")
```


This shows the mean and median of the data with the missing values replaced
```{r}

kbl(replacedMeanPerDay) %>% 
    kable_styling()
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
isDateWeekend <- 
    weekdays(dfInputMissing$dateTime) %in% c("Saturday","Sunday")
dfInputMissing$is.weekend <- 
    ifelse(isDateWeekend, "weekend","weekday")

dfInputMissing$is.weekend <- as.factor(dfInputMissing$is.weekend)

byWeekendSummary <- dfInputMissing %>%
    group_by(is.weekend, interval) %>% 
    summarise(mean.steps = mean(steps, na.rm = T),
              median.steps = median(steps, na.rm = T))

ggplot(data = byWeekendSummary, 
       mapping = aes(interval, mean.steps)) +
    geom_line()+
    facet_grid(is.weekend~.)
```




